---
title: "STAT 331 Portfolio"
author: "Ryan Chan"
format: 
  html: 
    self-contained: true
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an \_\_.

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv`

```{r}
#| label: wd-1-csv

# Source: Lab 2, Question 1
surveys = read_csv(here::here("data", "surveys.csv"),
                   show_col_types = FALSE)

```

-   `xlsx`

```{r}
#| label: wd-1-xlsx

# Source: Practice Activity 4
military <- read_xlsx(here::here("data",
                                 "gov_spending_per_capita.xlsx"),
                      sheet = "Share of Govt. spending",
                      skip = 7,
                      n_max = 191)

```

-   `txt`

```{r}
#| label: wd-1-txt

# Source: Check-in 2.3
ages_tab <- read_table(file = here::here("Week 2",
                                         "Check-ins",
                                         "Ages_Data",
                                         "ages_tab.txt"))

```

**WD-2: I can select necessary columns from a dataset.**

```{r}
#| label: wd-2

# Source: Challenge 3, Question 1
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(
    SET_level = ifelse(SET_score_avg >= 4,
                       "excellent", 
                       "standard"),
    sen_level = ifelse(seniority <= 4, 
                       "junior", 
                       "senior")
  ) |>
  select(
    course_id,
    SET_level,
    sen_level
)

```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   numeric

```{r}
#| label: wd-3-numeric

# Source: Lab 3, Question 9
teacher_evals_clean |>
  group_by(course_id, teacher_id) |>
  summarise(question_count = n_distinct(question_no)) |>
  filter(question_count == 9)

```

-   character -- specifically a string (example must use functions from **stringr**)

```{r}
#| label: wd-3-string

# Source: Lab 5, Finding Suspect
person |>
  inner_join(get_fit_now_member |>
               rename(membership_id = id),
             join_by(id == person_id)) |>
  left_join(get_fit_now_check_in,
            join_by(membership_id == membership_id)) |>
  left_join(drivers_license,
            join_by(license_id == id)) |>
  filter(ymd(check_in_date) == ymd(20180109),
         gender == "male",
         membership_status == "gold",
         str_detect(membership_id, "^48Z"),
         str_detect(plate_number, "H42W"))

```

-   factor

```{r}
#| label: wd-3-factor

# Source: Lab 4, Question 2
ca_childcare <- counties |>
  inner_join(childcare_costs,
             by = join_by(county_fips_code)) |>
  filter(state_name == "California")

```

-   date (example must use functions from **lubridate**)

```{r}
#| label: wd-3-date

# modify: I would recommend first renaming the id column of the person data **and then** piping into inner_join(). It is hard to understand what data are being joined when you are nesting functions inside of functions.

# Lab 5, Step 5
person |>
  inner_join(drivers_license,
             join_by(license_id == id)) |>
  inner_join(facebook_event_checkin,
             join_by(id == person_id)) |>
  filter(gender == "female",
         hair_color == "red",
         height %in% c(65:67),
         car_make == "Tesla",
         car_model == "Model S",
         event_name == "SQL Symphony Concert",
         month(ymd(date)) == 12,
         year(ymd(date)) == 2017) |>
  group_by(id) |>
  summarize(concert_attendance = n()) |>
  filter(concert_attendance == 3)

```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   numeric (using `as.numeric()` is not sufficient)

```{r}
#| label: wd-4-numeric

```

-   character -- specifically a string (example must use functions from **stringr**)

```{r}
#| label: wd-4-string

# Source: Lab 4, Question 3
ca_childcare <- ca_childcare |> 
  mutate(county_name = str_remove(county_name, " County"),
         region = fct_collapse(county_name,
                               "Superior California" = c("Butte", "Colusa", "El Dorado", "Glenn", "Lassen", "Modoc", "Nevada", "Placer", "Plumas", "Sacramento", "Shasta", "Sierra", "Siskiyou", "Sutter", "Tehama", "Yolo", "Yuba"),
                               "North Coast" = c("Del Norte", "Humboldt", "Lake", "Mendocino", "Napa", "Sonoma", "Trinity"),
                               "San Francisco Bay Area" = c("Alameda", "Contra Costa", "Marin", "San Francisco", "San Mateo", "Santa Clara", "Solano"),
                               "Northern San Joaquin Valley" = c("Alpine", "Amador", "Calaveras", "Madera", "Mariposa", "Merced", "Mono", "San Joaquin", "Stanislaus", "Tuolumne"),
                               "Central Coast" = c("Monterey", "San Benito", "San Luis Obispo", "Santa Barbara", "Santa Cruz", "Ventura"),
                               "Southern San Joaquin Valley" = c("Fresno", "Inyo", "Kern", "Kings", "Tulare"),
                               "Inland Empire" = c("Riverside", "San Bernardino"),
                               "Los Angeles County" = c("Los Angeles"),
                               "Orange County" = c("Orange"),
                               "San Diego-Imperial" = c("San Diego", "Imperial")))

```

-   factor (example must use functions from **forcats**)

```{r}
#| label: wd-4-factor

# Source: Lab 4, Question 3
ca_childcare <- ca_childcare |> 
  mutate(county_name = str_remove(county_name, " County"),
         region = fct_collapse(county_name,
                               "Superior California" = c("Butte", "Colusa", "El Dorado", "Glenn", "Lassen", "Modoc", "Nevada", "Placer", "Plumas", "Sacramento", "Shasta", "Sierra", "Siskiyou", "Sutter", "Tehama", "Yolo", "Yuba"),
                               "North Coast" = c("Del Norte", "Humboldt", "Lake", "Mendocino", "Napa", "Sonoma", "Trinity"),
                               "San Francisco Bay Area" = c("Alameda", "Contra Costa", "Marin", "San Francisco", "San Mateo", "Santa Clara", "Solano"),
                               "Northern San Joaquin Valley" = c("Alpine", "Amador", "Calaveras", "Madera", "Mariposa", "Merced", "Mono", "San Joaquin", "Stanislaus", "Tuolumne"),
                               "Central Coast" = c("Monterey", "San Benito", "San Luis Obispo", "Santa Barbara", "Santa Cruz", "Ventura"),
                               "Southern San Joaquin Valley" = c("Fresno", "Inyo", "Kern", "Kings", "Tulare"),
                               "Inland Empire" = c("Riverside", "San Bernardino"),
                               "Los Angeles County" = c("Los Angeles"),
                               "Orange County" = c("Orange"),
                               "San Diego-Imperial" = c("San Diego", "Imperial")))
```

-   date (example must use functions from **lubridate**)

```{r}
#| label: wd-4-date

# use same as above for WD-3 - date, make the inner_join a mutate()

```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()`

```{r}
#| label: wd-5-left

# Source: Lab 5, Step 6
person |>
  left_join(interview,
            join_by(id == person_id)) |>
  filter(id == 99716) |>
  select(name,
         transcript)

```

-   `right_join()`

```{r}
#| label: wd-5-right

```

-   `inner_join()`

```{r}
#| label: wd-5-inner

```

-   `full_join()`

```{r}
#| label: wd-5-full

```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi

```

-   `anti_join()`

```{r}
#| label: wd-6-anti

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()`

```{r}
#| label: wd-7-long

# Source: Lab 4, Question 6
ca_childcare |>
  filter(!is.na(mc_infant),
         !is.na(mc_toddler),
         !is.na(mc_preschool)) |>
  pivot_longer(cols = c(mc_infant,
                        mc_toddler,
                        mc_preschool),
               names_to = "age_group",
               values_to = "weekly_price") |>
  mutate(age_group = factor(age_group,
                            levels = c("mc_infant", "mc_toddler", "mc_preschool"),
                            labels = c("Infant", "Toddler", "Preschool")),
         region = factor(region,
                         levels = region_order)) |>
  # create the plot
  ggplot(mapping = aes(x = study_year,
                       y = weekly_price,
                       color = region)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess") +
  facet_wrap(~ age_group) +
  labs(x = "Study Year",
       y = "",
       subtitle = "Weekly Median Price for Center-Based Childcare ($)",
       color = "California Region") +
  scale_x_continuous(breaks = seq(2008,
                                  2018,
                                  by = 2)) +
  scale_y_continuous(limits = c(100, 500),
                     labels = label_currency(prefix = "$", # adding $ to y-axis labels
                                             accuracy = 1)) +
  theme(legend.position = "right",
        strip.background = element_rect(color = "black"),
        panel.border = element_rect(color = "black",
                                    fill = NA), # black border around each plot
        aspect.ratio = 1.1,
        axis.text = element_text(size = 6)
  )

```

-   `pivot_wider()`

```{r}
#| label: wd-7-wide

# Source: Lab 4, Question 4
ca_childcare |>
  group_by(region, study_year) |>
  filter(study_year %in% c(2008, 2018)) |>
  summarize(median_mhi = median(mhi_2018,
                                na.rm = TRUE)) |>
  pivot_wider(names_from = study_year,
              values_from = median_mhi) |>
  arrange(desc(`2018`))

```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

I've done this in the following provided assignments:

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting

```{r}
#| label: r-2-1

```

-   Example of **dplyr** pipeline

```{r}
#| label: r-2-2

```

-   Example of function formatting

```{r}
#| label: r-2-3

```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example -- any context

```{r}
#| label: r-3-example

```

-   Example of function stops

```{r}
#| label: r-3-function-stops

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   at least two numeric variables

```{r}
#| label: dvs-1-num

```

-   at least one numeric variable and one categorical variable

```{r}
#| label: dvs-2-num-cat

```

-   at least two categorical variables

```{r}
#| label: dvs-2-cat

```

-   dates (timeseries plot)

```{r}
#| label: dvs-2-date

```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can ensure people don't tilt their head

```{r}
#| label: dvs-2-1

```

-   I can modify the text in my plot to be more readable

```{r}
#| label: dvs-2-2

```

-   I can reorder my legend to align with the colors in my plot

```{r}
#| label: dvs-2-3

```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors

```{r}
#| label: dvs-3-1

```

-   I can use annotations

```{r}
#| label: dvs-3-2

```

-   I can be creative...

```{r}
#| label: dvs-3-3

```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()`

```{r}
#| label: dvs-4-summarize

```

-   Example using `across()`

```{r}
#| label: dvs-4-across

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r}
#| label: dvs-5-1

```

-   Example 2

```{r}
#| label: dvs-5-2

```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   Example 1

```{r}
#| label: dvs-6-1

```

-   Example 2

```{r}
#| label: dvs-6-2

```

**DVS-7: I show creativity in my tables.**

-   Example 1

```{r}
#| label: dvs-7-1

```

-   Example 2

```{r}
#| label: dvs-7-2

```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls)

```{r}
#| label: pe-1-one-call

```

-   `across()`

```{r}
#| label: pe-1-across

```

-   `map()` functions

```{r}
#| label: pe-1-map-1

```

**PE-2: I can write functions to reduce repetition in my code.**

-   Function that operates on vectors

```{r}
#| label: pe-2-1

```

-   Function that operates on data frames

```{r}
#| label: pe-2-2

```

**PE-3:I can use iteration to reduce repetition in my code.**

-   `across()`

```{r}
#| label: pe-3-across

```

-   `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1

```

-   `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated

```{r}
#| label: pe-4-1

```

-   I can connect a data wrangling pipeline into a `ggplot()`

```{r}
#| label: pe-4-2

```

## Data Simulation & Statisical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r}
#| label: dsm-1-1

```

-   Example 2

```{r}
#| label: dsm-1-2

```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1

```{r}
#| label: dsm-2-1

```

-   Example 2

```{r}
#| label: dsm-2-2

```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->
